---
title: "Introduction to `ggplot2`"
editor: visual
filters:
- webr
---

When I first started out with R, I only ever used the base plotting system. This was primarily because once I saw an example of a plot made with `ggplot` I was a little bit scared.

After awhile I realized that `ggplot` was the way to go when it came to graphics. The goal of this is to hopefully assuage you to do the same!

First, we will generate some data. I teach so we will use students and scores on a test and then look at how the students performed.

Some of my colleagues use a 4 test grading scheme where each test is worth 25%. Let's create a sample dataset with a class of 40 students and varying scores for each student as well as an average for each student.

```{r}
library(ggplot2)
library(forcats)
library(huxtable)
library(stringr)
library(dplyr,warn.conflicts = F)

test_data <- 
  data.frame(
    Student = paste0("Student",1:40),
    Exam_1= runif(40,80,85) |> round(0),
    Exam_2= runif(40,75,100) |> round(0),
    Exam_3= runif(40,70,90) |> round(0),
    Exam_4= runif(40,60,100) |> round(0)
  )

test_data |> 
  select(!Student) |> 
  summarize_all(mean) |> 
  hux() |> 
  theme_article()
```

Let's make a plot of the average score on each exam using the base plotting system.

```{r}

test_means <- 
  test_data |> 
  select(!Student) |> 
  summarize_all(mean) |> 
  tidyr::pivot_longer(cols = everything(),
                      names_to = "Exam",
                      values_to = "Average")

barplot(
  test_means$Average,
    names.arg = 
      test_means$Exam,
    col=c("darkblue","darkgreen","darkorange"
    ,"darkred"),
        density = 40,
        ylim = c(0,100),
        xlab="Exam Number",
        ylab="Score",
        main="Exam Score Averages")
```

This does not look bad, but we could probably do better with `ggplot`.

The first thing to note is the syntax that `ggplot` uses.

`ggplot(data,aes(x,y)) +   geom_() +   labs() +   theme_classic() +   theme()`

You will see here that each line is followed by a `+` which indicates that a new layer is being added. Let's try this out using our dataframe. I will do it the way you will see it most places, and then afterwards only do it the shortened way:

```{r}
# Use the means we created before and add two line numbers
# The x-axis will be the test number, y will be the average 
# The color will change based on value of the average

ggplot(data = test_means,
       aes(x = Exam, y = Average,fill = Exam)) +
  # Tells R to plot the value associated with the x-axis
  geom_bar(stat="identity",
           show.legend = FALSE,
           aes(fill = 
                 ifelse(Average == max(Average),
                  "darkgreen","grey"))) +
  # A nice theme 
  theme_minimal() +
  # Specifies x and y axes labels and title for plot
  labs(x = "\nExam Number",
       y = "Score\n",
       title = "Exam Score Averages") +
  ylim(0,100) + 
  scale_fill_identity()
```

I know that there is a *lot* going on in the example, but it kind of looks nice, right?

Let's go back to the example and add some filtering. Let's make a new column that determines whether or not a student passed or failed the test.

When we first made the dataset we included a fifth column, FinalGrade that represents the students overall grade. We will then make a series of statements that will determine what the letter grade is, and then plot the result!

We will be using the `|>` or pipe operator to pass the contents of `test_data` into our arranging.

We will first use the `mutate` function which creates a new column based on the arguments you pass to it.

We will also be using `case_when` which acts as an `ifelse` statement but with multiple conditionals.

-   The basic structure is: `case_when(variable=this~make it this)`

```{r}

# Create a new column for final grade

final_grade <- 
  test_data |> 
  group_by(Student) |> 
  rowwise() |> 
  mutate(
    FinalGrade = mean(c_across(Exam_1:Exam_4))
  ) |> 
  ungroup() |> 
  mutate(
    LetterGrade=
      case_when(
      FinalGrade>=95~"A",
      FinalGrade>=90~"A-",
      FinalGrade>=87~"B+",
      FinalGrade>=84~"B",
      FinalGrade>=80~"B-",
      FinalGrade>=74~"C+",
      FinalGrade>=70~"C",
      FinalGrade>=67~"D+",
      FinalGrade>=64~"D",
      FinalGrade>=60~"D-",
      FinalGrade<60~"F")
    ) |> 
  group_by(LetterGrade) |> 
  count()
```

Our students didn't do particularly well, but that is to be expected when we control the range their grades can be in!

Next, we will graph the data using `ggplot`, but this time we will pass the dataframe directly into the plot and remove the  `x=` and `y=` declarations

When we graph the results the letter grades will be backwards so we will need to use `fct_rev` to reverse them.

```{r}
final_grade %>% 
  # Place Grades on X axis and count on Y axis
  # Color based on Letter Grade
  ggplot(aes(fct_rev(LetterGrade),n,fill=LetterGrade)) +
    geom_bar(stat="identity") +
    theme_minimal(base_size=12) +
    labs(x="",y="Count",
         title="Letter Grade Frequency") +
    # Flip the axes for better readability
    coord_flip() +
    # Remove the legend and the axis ticks
    theme(legend.position ="none",
          axis.ticks = element_blank()) +
    # Add the count to the bars
    geom_text(aes(label=n,
                  hjust=-.08)) +
  theme(
    axis.text.x = element_blank()
  )
```

### Scatterplots

When I was first learning R in graduate school the class was given a task to plot points but in different colors depending on what gender the points came from. In base plotting this requires the use of not only the `plot` function but also the `points` function. It is entirely doable, but `ggplot` just handles it better!

Let's generate some data where we have hypothetical participants take an experiment where their reaction time is measured as well as their mood is measured.

```{r}
scatter <- 
  data.frame(
    Group = sample(c("G1","G2"),100,replace = T),
    RT = rnorm(100,800,150) |> round(2),
    TestScore = c(runif(50,60,100) |> round(2),
                runif(50,40,90) |> round(2)
                )
  )
scatter |> 
  head() |> 
  hux() |> 
  theme_article()
```

Plot Sex against RT and have the colors be linked to the Group of the participant (point).

```{r}
scatter %>% 
  ggplot(aes(TestScore,RT,color = Group)) +
    geom_point(aes(shape = Group)) +
    theme_minimal() +
    labs(x="\nTest Score",
         y="RT(ms)\n",
         title="RT and Test Scores",
         subtitle = "Participants had quicker reaction time in Group 2 compared to Group 1\n") + 
  theme(
    plot.title = element_text(size = 20, 
                              face = "bold"),
    plot.subtitle = element_text(face = "italic")
  )
```

No pattern is emerging from the (fake) data, but we can see a distinction between the different points.

### Heatmaps

One of my favorite `geom`'s to use is `geom_tile` which creates a heatmap style plot.

Let's create a fake TV show with fake episode titles and fake episode ratings.

```{r}
Dave_TV <- 
  data.frame(
    Season = rep(paste0("S",1:6),each=12),
    Episode = as.factor(rep(1:12,6)),
    Title = str_to_title(paste0("The ",
                              sample(words,72)," ",
                              sample(words,72))),
    Ratings = c(runif(60,5.5,9),runif(12,7,9)) |> round(0),
    stringsAsFactors = F)

Dave_TV |> head() |> hux() |> theme_article()
```

Firstly, I want to take a moment to see how truly excellent some of the episode titles are:

```{r}
title_samp <- sample(Dave_TV$Title,10)
title_samp
```

Anyways, now that our data is in the correct format we can plot it.

```{r}
Dave_TV %>% 
  ggplot(aes(Season,Episode,fill = Ratings)) +
    geom_tile() +
    theme_minimal() +
    labs(x = "\n") +
    geom_text(aes(label = Ratings)) +
    theme(axis.ticks = element_blank())
```

There could be an entirely separate tutorial on just color choices in `ggplot2`.

Here is one last iteration that uses conditional values to color each episode or tile.

```{r}
cols=c("Bad"="red2",
       "Garbage" ="darkblue",
       "Great" ="darkgreen",
       "Regular"="darkorange", 
       "Good" ="gold")

Dave_TV %>% 
  mutate(Quality=case_when(
      Ratings<5.0 ~ "Garbage",
      Ratings<6.5 ~ "Bad",
      Ratings<7.5 ~ "Regular",
      Ratings<8.5 ~ "Good",
      Ratings<10 ~ "Great")) %>% 
  ggplot(aes(Season,Episode,fill=Quality)) +
    geom_tile(color="black",size=.085) +
    theme_minimal() +
    geom_text(aes(label=Ratings)) +
    theme(axis.ticks = element_blank()) +
    scale_fill_manual(values=cols) +
    labs(x="",y="",
        title="Dave TV Episode Ratings",
        caption="Data from IAmDb)") +
  theme(
    plot.caption = element_text(face = "italic")
  )
```

So there you have it, a very *brief* introduction to `ggplot`. As my own projects expand, I will include new sections or add more examples here.
