---
title: "Webscraping"
editor: visual
---

#### Using `dplyr` and `ggplot` for Visualization

```{r}
#| message: false
#| warning: false
library(ggplot2)
library(stringr)
library(forcats)
library(gganimate)
library(dplyr)
library(tidytext)
library(plotly)
library(huxtable)
```

I am still getting used to the idea of a few of these functions so if this seems all over the place, there is a reason for that!

I came across [this](https://www.reddit.com/r/dataisbeautiful/comments/fw4iv0/oc_rating_of_simpsons_episodes_according_to_imdb/) post on [r/dataisbeautiful](http://davidabrocker.com/reddit.com/r/dataisbeautiful) and I was instantly convinced to give it a shot!

The method I used was a little bit convoluted and hacky. What I am going to show here is how to read data off of a website and then show you how the data looks, and how to plot it in the way we would like.

Further down the line when I am more comfortable with the complex, I can make a more detailed tutorial.

For now, we will be downloading the file hosted [here](https://raw.githubusercontent.com/DavidBrocker/personal/master/files/DoctorWho_Ratings.csv)

```{r}
# First read in the csv 
doctorwho <- 
  read.csv("doctorwho_ratings.csv")
```

```{r}
head(doctorwho)
```

From our data set we can see that we have information on the following:

-   Season Number

-   Episode Number

-   Episode Title

-   IMDb User Rating

-   Categorical Rating

We can do a few things with this data. First, lets see which doctor has the most episodes.

```{r}
doctorwho %>% 
  group_by(Doctor) %>% 
  count(sort=T) |> 
  hux() |> 
  theme_article()
```

The eleventh doctor wins out here!

Next it might be fun to see which season had the highest rating and then graph that.

```{r}
# These are the custom colors we will be using for our ratings
dwcolor <- 
  c("Bad"="red2", "Garbage" ="dodgerblue2", 
    "Great" ="greenyellow", "Regular"="darkorange1", 
    "Good" ="gold1")

doctorwho |> 
  group_by(Season) |> 
  ggplot(aes(as.factor(Season),Ratings)) + 
  geom_point(aes(color = Quality),
             alpha = .2) +
  stat_summary(
    fun = "mean",
    geom = "point"
  ) + 
  theme_minimal() +
  theme(legend.position = "none",
        axis.ticks = element_blank()) +
  labs(x="\nSeason",y="Rating\n",
       title="Doctor Who Average Season Rating\n",
       caption="(based on IMDb User Ratings)") +
  theme(
    plot.caption = 
      element_text(
      face = "italic"),
    plot.title = 
      element_text(
      face = "bold")
  ) +
  #scale_color_manual(values = dwcolor)
  scale_color_discrete()

```

The takeaway from this is that the last two seasons were rated as the worst with Season 4 being rated as the best.

Another question might be which Doctor has the highest rated episodes?

Luckily we have a column with categorical information that tells us what ratings correspond to "Great".

Here is the system used: \< 5.0 = Garbage \< 6.5 = Bad \< 7.5 = Regular \< 8.5 = Good \< 10 = Great

So we could either `filter(Ratings>8.5)` or we could `filter(Quality=="Great")`.

Grouping by Doctor ensures that the information of who the Doctor was that received that rating.

```{r}
doctorwho %>% 
  group_by(Doctor,Ratings) %>% 
  filter(Quality=="Great") %>% 
  group_by(Doctor) %>% 
  count() %>% 
  ggplot(aes(fct_reorder(Doctor,n),n,fill=Doctor,label = n)) +
    geom_bar(stat="identity") +
    theme_minimal()+
  coord_flip() +
    theme(
      axis.ticks = element_blank(),
      legend.position = "none",
      plot.caption = element_text(face = "italic"),
      axis.text.x = element_blank(),
      axis.text.y = element_text(face = "bold")
      ) +
    labs(x="",y="",
         title="Number of 'Great' Episodes: Per Doctor\n",
         captions="(based on IMDb User Ratings)") +
  scale_fill_viridis_d(alpha = .6) +
  geom_label(color = "black") 
```

To be honest, no surprise here either!

Lastly, let's construct a heatmap of all of the episodes of Doctor Who.

=

```{r}
# Plotly is a package add-on that will make our graph interactive
# This is how it can be specially formatted
doctorwho_pltly <- 
  doctorwho %>% 
  mutate(text=
           paste0("Season: ", Season, "\n",
                  "Episode: ", EpisodeNumber,"\n",
                  "Title: ", Title, "\n",
                  "Doctor: ",Doctor))

library(ggtext)
p <- 
  doctorwho %>% 
  ggplot(aes(factor(Season),EpisodeNumber,fill=Quality)) +
    geom_tile(colour="black")+
    geom_text(aes(label=Ratings))+
    scale_fill_manual(values=dwcolor)+
  theme_minimal()+
  theme(axis.ticks = element_blank(),
        plot.subtitle = element_markdown()) +
    labs(x="Season\n",
         y="\nEpisode",
         title="Doctor Who IMDb Ratings: 2005-2020",
         subtitle = "Ratings of the <span style='color:#0072B2;'>good</span>
         and <span style='color:#009E73;'>bad</span>")
  
ggplotly(p)
```
